---
title: "BottomsUp"
author: "David Stroud"
date: "2/12/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Cory Test

### Nice job  ~ dstroud

/* Cory's sad first attempt */

##Merge
```{r merge, include=TRUE}
Beers <- read.csv("C:/Users/william.thigpen/Dropbox/Doing Data Science/Case Study 1/Beers.csv")
Breweries <- read.csv("C:/Users/william.thigpen/Dropbox/Doing Data Science/Case Study 1/Breweries.csv")
names(Beers) <- c("Name", "Beer_ID", "ABV", "IBU", "Brew_ID")
brewset <- merge(Beers, Breweries, by="Brew_ID", all=FALSE)
```

##Question 1
###How many breweries are present in each state?
```{r q1, include=TRUE}
summary(brewset$State)
```

##Question 2
###Merge beer data with the breweries data. Print the first 6 observations and the last six observations to check the merged file.
```{r q2, include=TRUE}
head(brewset, 6)
tail(brewset, 6)
```

##Question 3
###Report the number of NA's in each column.
```{r q3, include=TRUE}
sapply(X = brewset, FUN = function(x) sum(is.na(x)))
```

##Question 4
###Compute the median alcohol content and international bitterness unit for each state. Plot a bar chart to compare.
```{r q4, include=TRUE}
ABVmedians <- aggregate(brewset$ABV, list(brewset$State), median, na.rm=TRUE)
barplot(`colnames<-`(t(ABVmedians[-1]), ABVmedians[,1]), beside=TRUE, legend.text = TRUE, col = c("black", "grey"), cex.names=.6)
IBUmedians <- aggregate(brewset$IBU, list(brewset$State), median, na.rm=TRUE)
barplot(`colnames<-`(t(IBUmedians[-1]), IBUmedians[,1]), beside=TRUE, legend.text = TRUE, col = c("black", "grey"), cex.names=.6)
```

/* This could surely be done better */
##Question 5
###Which state has the maximum alcoholic (ABV) beer? Which state has the most bitter (IBU) beer?
```{r q5, include=TRUE}
aggregate(brewset$ABV, list(brewset$State), max, na.rm=TRUE)
```

##Question 6
###Summary statistics for the ABV variable.
```{r q6, include=TRUE}
summary(brewset$ABV)
```

/* This is a sadly coded plot */
##Question 7 
###Is there an apparent relationship between the bitterness of the beer and its alcoholic content? Draw a scatter plot.
```{r q7, include=TRUE}
plot(brewset$IBU, brewset$ABV, na.rm=TRUE)
```
